### 2024-11-02 manuscript.tex
Update on Overleaf.

The main aim of this study is to evaluate the GSRM version of ICON, developed
jointly by nextGEMS, Deutscher Wetterdienst, Max-Planck-Institute for
Meteorology, Deutsches Klimarechenzentrum (DKRZ), Karlsruhe Institute of
Technology, and the Center for Climate Systems Modeling.  Previous studies have
identified substantial large-scale biases in climate model clouds over the SO,
affecting sea surface temperature and the Earth’s albedo. Our aim is to
quantify how well the GSRM ICON is simulating clouds in this region,
particularly in light of the fact that subgrid-scale clouds and convection are
not parameterized in this model. This region is mostly dominated by boundary
layer clouds generated by shallow convection, and these are problematic to
observe by spaceborne lidars and radars, which are affected by attenuation by
overlapping and thick clouds and ground clutter, respectively.  Specifically,
the radar on CloudSat and lidar on CALIPSO (neither of them operational any
more) are affected by the abovementioned issues, resulting in a strong
underestimation of cloud occurrence below 2 km relative to ground-based lidar
observations \citep{mcerlich2021}.  This, in turn, can lead to systematic
biases in low clouds in climate models, which are frequently evaluated against
CloudSat--CALIPSO products. Reanalyses can also suffer from cloud biases, as
these are usually parametrised in their atmospheric component, and also in
regions where input observations are sparse.  This makes them a problematic
reference for clouds over the SO, and any biases relative to a reanalysis
should be interpreted with caution. Instead, we chose to use a large set of
ship-based observations conducted with ceilometers and lidars on board of the
RV \emph{Polarstern} and other voyages and stations as a reference for the
model evaluation.

@@

Together, we analysed data from 31 voyages of RV \emph{Polarstern}, resupply
vessel (RSV) \emph{Aurora Australis}, RV \emph{Tangaroa}, RV \emph{Nathaniel B.
Palmer}, Her Majesty's New Zealand Ship (HMNZS) \emph{Wellington} and one
sub-Antarctic station (Macquarie Island) in the SO south of 40°S between 2010
and 2021. Fig. \ref{fig:map} shows a map of the voyages, Table
\ref{tab:voyages} list the voyages, campaigns, and stations, and Table
\ref{tab:voyage-references} lists references where available. Altogether, the
voyages and station dataset comprised 2421 days of data south of 40°S, but the
availability of ceilometer data was slightly smaller due to gaps in
measurements.

@@

The data sources contained ceilometer observations captured by the Vaisala CL51
operating at a wavelenght of 910 nm, the Vaisala CT25K operating at 905 nm, and
the Lufft CHM 15k operating at 1064 nm, described in detail below (Sections
\ref{sec:cl51} and \ref{sec:chm15k}). A ceilometer is a low-power near-infrared
vertically pointing lidar with the purpose of measuring the cloud base, but it
also measures the full vertical structure of clouds as long as the laser signal
is not attenuated by thick clouds, which can be used to infer additional
information such as a cloud mask and cloud occurrence by height.

@@

The Vaisala CL51 (photo in Fig. \ref{fig:map}d) and CT25K are ceilometers
operating at a near-infrared wavelength of 910 nm and 905 nm, respectively.
CL51 can also be configured to emulate the Vaisala CL31. The maximum range is
15.4 km (CL51), 7.7 km (CL31 emulation mode with 5 m vertical resolution), and
7.5 km (CT25K). The vertical resolution is 10 m (5 m configurable) in CL51 and
30 m in CT25K. The sampling (temporal) resolution is configurable, and in our
datasets it is about 6 s for CL51 on AA15‐16, 16 s for CT25K on MARCUS and
MICRE, 36 s for CL51 on RV \emph{Polarstern}, and about 2.37 s for CL51 with
CL31 emulation on TAN1502. The wavelength of 910 nm is affected by water vapour
absorption of about 20\% in the mid-latitudes \citep{wiegner2015,wiegner2019},
but we do not expect this to be a significant issue as explained in
\cite{kuma2021}.  The instrument data files containing raw uncalibrated
backscatter were first converted to Network Common Data Form (NetCDF) with
cl2nc (\url{https://github.com/peterkuma/cl2nc}) and then processed with the
ALCF (Section \ref{sec:alcf}) to produce absolutely calibrated attenuated
volume backscattering coefficient (AVBC), cloud mask, cloud occurrence by
height, and the total cloud fraction. Because the CT25K uses a very similar
wavelength to CL51, equivalent calculations as for CL51 were done assuming a
wavelength of 910 nm. The Vaisala CL51 and CT25K instruments were used on most
of the voyages and stations analysed here. Fig.  \ref{fig:example}a shows an
example of AVBC derived from the CL51 instrument data.

\begin{figure}[b!]
\centering
\includegraphics[width=\textwidth]{img/example.png}
\caption{
An example of attenuated volume backscattering coefficient (AVBC) \textbf{(a)}
measured by CL51 during 24 hours on the PS81/3 voyage and \textbf{(b)} an
equivalent AVBC simulated with the ALCF from ERA5 data during the same time
period. Shown is also the cloud mask determined by the ALCF.
}
\label{fig:example}
\end{figure}

@@

The Lufft CHM 15k (photo in Fig. \ref{fig:map}c) is a ceilometer operating at a
near-infrared wavelength of 1064 nm. The maximum range is 15.4 km, the vertical
resolution is 5 m in the near range (up to 150 m) and 15 m above, the sampling
(temporal) resolution is 2 s, and the number of vertical levels is 1024.
NetCDF files containing uncalibrated backscatter produced by the instrument
were processed with the ALCF (Section \ref{sec:alcf}) to produce AVBC, cloud
mask, cloud occurrence by height, and the total cloud fraction. The CHM 15k was
used on four voyages (HMNZSW16, TAN1702, TAN1802, and NBP1704).

@@

The Automatic Lidar and Ceilometer Framework (ALCF) is a ground-based lidar
simulator and a tool for processing observed lidar data, supporting various
instruments and models \citep{kuma2021}. It performs radiative transfer
calculations to derive equivalent lidar AVBC in an atmospheric model, which can
then be compared with observed AVBC. For this purpose, it takes atmospheric
fields of cloud fraction, liquid and ice mass mixing ratio, temperature, and
pressure fields as an input and is run offline (on the model output rather than
inside the model code). The lidar simulator in the ALCF is based on the
instrument simulator Cloud Feedback Model Intercomparison Project (CFMIP)
Observation Simulator Package (COSP) \citep{bodas-salcedo2011}.  After AVBC is
calculated, a cloud mask, cloud occurrence by height, and the total cloud
fraction are determined. The ALCF has been used by several research teams for
model and reanalysis evaluation
\citep{kuma2020,kremser2021,guyot2022,pei2023,whitehead2023,mcdonald2024}.

>>>>

The main aim of this study is to evaluate the GSRM version of ICON, developed
jointly by nextGEMS, Deutscher Wetterdienst, Max-Planck-Institute for
Meteorology, Deutsches Klimarechenzentrum (DKRZ), Karlsruhe Institute of
Technology, and the Center for Climate Systems Modeling.  Previous studies have
identified substantial large-scale biases in climate model clouds over the SO,
affecting sea surface temperature and the Earth’s albedo. Our aim is to
quantify how well the GSRM ICON is simulating clouds in this region,
particularly in light of the fact that subgrid-scale clouds and convection are
not parameterized in this model. This region is mostly dominated by boundary
layer clouds generated by shallow convection, and these are problematic to
observe by spaceborne lidars and radars, which are affected by attenuation by
overlapping and thick clouds and ground clutter, respectively.  Specifically,
the radar on CloudSat and lidar on CALIPSO (neither of which are now operational) are affected by the abovementioned issues, resulting in a strong
underestimation of cloud occurrence below 2 km relative to ground-based lidar
observations \citep{mcerlich2021}.  This, in turn, can lead to systematic
biases in low clouds in climate models, which are frequently evaluated against
CloudSat--CALIPSO products. Reanalyses can also suffer from cloud biases, as
these are usually parametrised in their atmospheric component, and also in
regions where input observations are sparse.  This makes them a problematic
reference for clouds over the SO, and any biases relative to a reanalysis
should be interpreted with caution. Instead, we chose to use a large set of
ship-based observations conducted with ceilometers and lidars on board of the
RV \emph{Polarstern} and other voyages and stations as a reference for the
model evaluation.

@@

Together, we analysed data from 31 voyages of RV \emph{Polarstern}, the resupply
vessel (RSV) \emph{Aurora Australis}, RV \emph{Tangaroa}, RV \emph{Nathaniel B.
Palmer}, Her Majesty's New Zealand Ship (HMNZS) \emph{Wellington} and one
sub-Antarctic station (Macquarie Island) in the SO south of 40°S between 2010
and 2021. Fig. \ref{fig:map} shows a map of the voyages, Table
\ref{tab:voyages} list the voyages, campaigns, and stations, and Table
\ref{tab:voyage-references} lists references where available. Altogether, the
voyages and station dataset comprised 2421 days of data south of 40°S, but the
availability of ceilometer data was slightly smaller due to gaps in
measurements.

@@

The data sources contained ceilometer observations captured by the Vaisala CL51
operating at a wavelenght of 910 nm, the Vaisala CT25K operating at 905 nm, and
the Lufft CHM 15k operating at 1064 nm, described in detail below (Sections
\ref{sec:cl51} and \ref{sec:chm15k}). A ceilometer is a low-power near-infrared
vertically pointing lidar principally designed to  measure cloud base, but they 
also measure the full vertical structure of clouds as long as the laser signal
is not attenuated by thick clouds, which can be used to infer additional
information such as a cloud mask and cloud occurrence by height.

@@

The Vaisala CL51 (photo in Fig. \ref{fig:map}d) and CT25K are ceilometers
operating at a near-infrared wavelength of 910 nm and 905 nm, respectively.
The CL51 can also be configured to emulate the Vaisala CL31. The maximum range is
15.4 km (CL51), 7.7 km (CL31 emulation mode with 5 m vertical resolution), and
7.5 km (CT25K). The vertical resolution is 10 m (5 m configurable) in CL51 and
30 m in CT25K observations. The sampling (temporal) resolution is configurable, and in our
datasets is approximately 6 s for CL51 on AA15‐16, 16 s for CT25K on MARCUS and
MICRE, 36 s for CL51 on RV \emph{Polarstern}, and about 2.37 s for CL51 with
CL31 emulation on TAN1502. The wavelength of 910 nm is affected by water vapour
absorption of about 20\% in the mid-latitudes \citep{wiegner2015,wiegner2019},
but we do not expect this to be a significant issue as explained in
\cite{kuma2021}.  The instrument data files containing raw uncalibrated
backscatter were first converted to Network Common Data Form (NetCDF) with
cl2nc (\url{https://github.com/peterkuma/cl2nc}) and then processed with the
ALCF (Section \ref{sec:alcf}) to produce absolutely calibrated attenuated
volume backscattering coefficient (AVBC), cloud mask, cloud occurrence by
height, and the total cloud fraction. Because the CT25K uses a very similar
wavelength to CL51, equivalent calculations as for CL51 were done assuming a
wavelength of 910 nm. The Vaisala CL51 and CT25K instruments were used on most
of the voyages and stations analysed here. Fig.  \ref{fig:example}a shows an
example of AVBC derived from the CL51 instrument data.

\begin{figure}[b!]
\centering
\includegraphics[width=\textwidth]{img/example.png}
\caption{
An example of attenuated volume backscattering coefficient (AVBC) \textbf{(a)}
measured by CL51 during 24 hours on the PS81/3 voyage and \textbf{(b)} an
equivalent AVBC simulated with the ALCF from ERA5 data during the same time
period. The red line identifies the cloud mask determined by the ALCF.
}
\label{fig:example}
\end{figure}

@@

The Lufft CHM 15k (photo in Fig. \ref{fig:map}c) is a ceilometer operating at a
near-infrared wavelength of 1064 nm. The maximum range is 15.4 km, the vertical
resolution is 5 m in the near range (up to 150 m) and 15 m above, the sampling
(temporal) resolution is 2 s, and the number of vertical levels is 1024.
NetCDF files containing uncalibrated backscatter produced by the instrument
were processed with the ALCF (Section \ref{sec:alcf}) to again produce AVBC, cloud
mask, cloud occurrence by height, and the total cloud fraction. The CHM 15k was
used on four voyages (HMNZSW16, TAN1702, TAN1802, and NBP1704).

@@

The Automatic Lidar and Ceilometer Framework (ALCF) is a ground-based lidar
simulator and a tool for processing observed lidar data, supporting various
instruments and models \citep{kuma2021}. It performs radiative transfer
calculations to derive equivalent lidar AVBC in an atmospheric model, which can
then be compared with observed AVBC. For this purpose, it takes the cloud fraction, liquid and ice mass mixing ratio, temperature, and
pressure model fields as an input and is run offline (on the model output rather than
inside the model code). The lidar simulator in the ALCF is based on the
instrument simulator Cloud Feedback Model Intercomparison Project (CFMIP)
Observation Simulator Package (COSP) \citep{bodas-salcedo2011}.  After AVBC is
calculated, a cloud mask, cloud occurrence by height, and the total cloud
fraction are determined. The ALCF has been used by several research teams for
model and reanalysis evaluation
\citep{kuma2020,kremser2021,guyot2022,pei2023,whitehead2023,mcdonald2024}.
